---
title: "BASA arthropods"
author: "Colleen Nell"
date: "9/25/2017"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, fig.width = 6, cache=TRUE)

setwd('/Users/colleennell/Dropbox/Projects/BASA')

library(devtools)
library(ggplot2)
library(dplyr)
library(data.table)
library(lme4) #LMM and max liklihood tests
library(afex) #for parametric bootstrap simulations of deviance stat in LMM
library(car)
library(lsmeans)
library(labdsv)
library(vegan)
library(cowplot)

source('BASA_block.r')#add column separating plants into spatial blocks
source('/Users/colleennell/Dropbox/rstats/theme_nell.R')
source('/Users/colleennell/Dropbox/Projects/BASA/basa_themes.R')

se<-function(x) sd(x, na.rm=TRUE)/sqrt(length(x))

set.seed(666)
```  

## BASA common garden data  
```{r, trait_in, include=FALSE}
##raw trait data
traits<-read.csv('traits_herb_terp_final.csv')%>%
  mutate(GENO = ifelse(SEX=='F', FAM-20, FAM))%>%
  mutate(RGR =  (log(Length_2011)-log(Length_2008))/(3), RGR = ifelse(Length_2011 > 0, RGR, NA), newRGR = log(Length_2011/4))%>%
  transform(ROW = ordered(ROW), COL = ordered(COL), FAM = factor(FAM), GENO = factor(GENO), BLOCK = as.factor(BLOCK))

##trait lsms
lsm.row<-read.csv('LSMS_GENO_ROW.csv')
lsm.row$lsmean<-ifelse(lsm.row$lsmean < 0, 1, lsm.row$lsmean)
lsm.row<-read.csv('LSMS_GENO_ROW.csv')
lsm.cast<-lsm.row%>%dcast(SEX+FAM~trait, value.var='lsmean')

##arthropods raw data
arths_2010<-read.csv('data/BASA_arth_data_2010.csv')[-233:-234,]
arths_2011<-read.csv('data/BASA_arth_data_2011.csv')[-245:-246,]
##create df with taxa names, groupings, abundance for 2010 and 2011
arthropods<-arths_2010%>%melt(id.vars='ID', variable.name='taxa', value.name='2010')%>%
  full_join(arths_2011%>%melt(id.vars='ID', variable.name='taxa', value.name='2011'), by=c('ID','taxa'))%>%
  left_join(read.csv('data/BASA_arth_data_FAMID.csv'), by=c('taxa'='Proposed.ID'))%>% ##join with data on arth trophic level etc
  filter(!taxa %in% c('X','Sample14','sum','Sum'), ID !='sum', ID != '')%>%
  mutate(`2010` = ifelse(is.na(`2010`), 0, `2010`),`2011` = ifelse(is.na(`2011`), 0, `2011`),
         total_arths = `2010` + `2011`)

```  

```{r, include=FALSE}
## Generate community dataframes for multivariate analyses  
#at the plant and genotype level, calculate bray-curtis dissimilarity on relative density of orders

#arth_comm_ind(trophic = c('arths','Herb','Pred','Omni','Detrit'), group = c('Order','Family','taxa',etc), dist, log=TRUE/FALSE, rel.abun=TRUE/FALSE)
# returns a community df with grouping variables, community matrix, log transformed matrix, relative abundance matrix, and dissimilarity matrix

##for all arthropods at the order level
arth.plant<-arth_comm_ind(trophic='arths', group = 'Order', dist='bray', log=FALSE)

##genotype arthropod lsms
#by order
arth.geno<-list(comm.df = read.csv('GENO_LSM_ARTH_ORD_COMM.csv')%>%left_join(lsm.cast%>%dplyr::select(FAM,SEX,RGR,flowers_all, SLA,CN,MT_sum,ST_sum,Per_H2O, leaf.toughness)), comm.mat = read.csv('GENO_LSM_ARTH_ORD_MAT.csv'))
arth.geno$comm.dist <-vegdist(arth.geno$comm.mat, method='bray')

```  
I calculated the relative density of arthropod Orders (`r length(colnames(arth.plant$comm.mat))`;`r colnames(arth.plant$comm.mat)`) and calcualted pairwise Bray-Curtis dissimilarities between plants.  
```{r}
library(RColorBrewer)
##stacked bar of arthropod compisiotn by genotypes
gen.ord<-cbind(arth.geno$comm.mat, arth.geno$comm.df%>%dplyr::select(FAM,SEX))%>%
  melt(id.vars=c('SEX','FAM'), variable.name='Order')%>%arrange(Order)

gen.ord%>%group_by(Order)%>%summarize(mean=mean(value))%>%arrange(desc(mean))
mycols = c(brewer.pal(3, 'Oranges'), brewer.pal(3, 'Greens'), 
           brewer.pal(3, 'Blues'), brewer.pal(6, 'PuRd'))
levels(gen.ord$Order)
gen.ord$reOrder<-reorder(gen.ord$Order, r)

ggplot(gen.ord, aes(x=interaction(SEX,FAM), y=value, fill=reorder(Order,value)))+
  geom_bar(stat='identity', aes(alpha=SEX))+theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())+
  labs(y='Relative density', x='Genotypes')+scale_fill_manual(values=mycols)+scale_alpha_manual(values=c(1,.6))
```
### PERMANOVA for genotype effect controlling for sex, block, water  
```{r}
arth.ad<-adonis2(arth.plant$comm.dist~FAM, data=arth.plant$comm.df, permutations=1000, strata='SEX')
arth.ad
##get r2 values
arth.ad.r<-adonis(arth.plant$comm.dist~FAM, data=arth.plant$comm.df, permutations=10000)
arth.ad.r

##FAM rsq= 0.09??
```  
Stepwise permutation tests in constrained ordination  
```{r}
ordistep()
```


```{r}
######
##independent permanova within each sex to get Rsq
##generate dataframes, distance matrices
male.df<-arth.plant$comm.df%>%filter(SEX == 'M')%>%dplyr::select(FAM,BLOCK,ROW,WATER,ACARINA:THYSANOPTERA)
female.df<-arth.plant$comm.df%>%filter(SEX == 'F')%>%dplyr::select(FAM,BLOCK,ROW,WATER,ACARINA:THYSANOPTERA)
male.rel<-male.df%>%dplyr::select(-FAM, -BLOCK, -ROW, -WATER)%>%mutate(total=rowSums(.,na.rm=TRUE))%>%mutate_all(funs(./total))%>%dplyr::select(-total)
female.rel<-female.df%>%dplyr::select(-FAM, -BLOCK, -ROW, -WATER)%>%mutate(total=rowSums(.,na.rm=TRUE))%>%mutate_all(funs(./total))%>%dplyr::select(-total)
male.dist<-vegdist(male.rel, method='bray')
female.dist<-vegdist(female.rel, method='bray')

male.ad<-adonis2(male.dist~FAM+WATER, data=male.df, permutations=10000, strata='BLOCK')
male.ad ##p=0.089
male.ad.r<-adonis(male.dist~FAM+WATER, data=male.df, permutations=10000)
male.ad.r
##dispersion?
beta<-betadisper(male.dist, male.df$FAM)
vegan::permutest(beta)
varpart(male.dist, ~BLOCK, ~WATER,~FAM,data=male.df)

female.ad<-adonis2(female.dist~FAM+WATER, data=female.df, permutations=10000, strata='BLOCK')
female.ad ##females p =0,0009
female.ad.r<-adonis(female.dist~FAM+WATER, data=female.df, permutations=10000)
female.ad.r
beta<-betadisper(female.dist, female.df$FAM)
vegan::permutest(beta)
varpart(female.dist, ~BLOCK, ~WATER, ~FAM,data=female.df)

```  
Permanova results: arthropod composition differs with genotype (stratafied by BLOCK). But, independent tests of male and female arthropod communities actually showed that arthropods were marginally different across male genotypes (P=0.087; Rsq = 0.18) where composition on female genotypes was different (P = 0.0001, Rsq = 0.259).  

For all of these permanovas, the Rsq does not account for BLOCK due to constraints of the test. However, results hardly differ with and without permutations stratified within BLOCK so probably not a big deal?  

### Genetic variation in arthropod communities   
Points show genotype centroids based on ordination of arthropod relative density (Bray-Curtis dissimilarity) for each plant, while controlling for plant sex (conditioned). Biplot arrows show linear trends between arthropod Orders and MDS axes (P < 0.05), with the relative lengths scaled to their Rsq.  

The community triplot shows arthropod composition constrained by plant traits, to maximize variation explained (dbRDA). Trait arrows show traits that has significant marginal effects (Type III) on composition, scaled to reflect their relative influence. Angles between arrows are indicative of correlations.   

#### Unconstrained ordination  
```{r, include=FALSE}
pco.plant<-capscale(arth.plant$comm.dist ~ 1 + Condition(arth.plant$comm.df$SEX), data=arth.plant$comm.df, comm = arth.plant$comm.mat)
#anova(pco.plant, permutations=1000)
head(summary(pco.plant))
###########
#plot centroids for each genotype, unconstrained ordination
plant.ind<-data.frame(ID = rownames(pco.plant$CA$u), SEX= arth.plant$comm.df$SEX, 
                        FAM=arth.plant$comm.df$FAM, WATER=arth.plant$comm.df$WATER, BLOCK=arth.plant$comm.df$BLOCK,
                        CAP1 = scores(pco.plant, scaling=3, correlation=TRUE)$sites[,1], 
                        CAP2 = scores(pco.plant, scaling=3, correlation=TRUE)$sites[,2])

##this si scaling 2 (species axes correlations)
plant.sp<-data.frame(label=rownames(scores(pco.plant, scaling=3, correlation=TRUE)$species), 
                       CAP1 = scores(pco.plant, scaling=3, correlation=TRUE)$species[,1], 
                       CAP2 = scores(pco.plant, scaling=3, correlation=TRUE)$species[,2])

##use envfit for fitted order arrows
#inertia is variance in species abundance
bt<-cbind(arth.plant$comm.df%>%dplyr::select(ID,SEX,FAM), arth.plant$comm.mat)
pco.env<-envfit(pco.plant, bt)

fam.c<-data.frame(FAM = rownames(pco.env$factors$centroids), MDS1 = pco.env$factors$centroids[,1], MDS2 = pco.env$factors$centroids[,2])
fam.center<-fam.c%>%filter(grepl('FAM', FAM))%>%
    mutate(FAM =gsub('FAM','', FAM))%>%
    left_join(arth.plant$comm.df%>%dplyr::select(FAM,SEX)%>%unique()%>%transform(FAM=as.character()))
fam.center$SEX<-ifelse(fam.center$FAM > 20, 'F', 'M')
sex.center<-fam.c%>%filter(FAM == 'SEXM' | FAM =='SEXF')
plant.arrow<-data.frame(order= rownames(pco.env$vectors$arrows), MDS1 = pco.env$vectors$arrows[,1], MDS2 = pco.env$vectors$arrows[,2], pval = pco.env$vectors$pvals, rsq = pco.env$vectors$r)%>%
  filter(pval <= 0.05, rsq > 0.1)
pco.env

##axis labels with var explained
xlabel<-paste0('MDS1 (', round(100*cumsum(eig / sum(eig))[[1]],2), '% of total variation)')
ylabel<-paste0('MDS2 (', round(100*(eig[[2]] / sum(eig)),2), '% of total variation)')

#anova(pco.plant, 'margins', permutaitons=10000)

```
```{r}
###build the plot
site.plot<-ggplot(fam.center, aes(x=MDS1, y=MDS2))+
    geom_jitter(aes(fill=SEX), shape=21,size=1.5, alpha=.8)+
    theme_nell()+
    scale_fill_manual(values=c('black','white'))+
    theme(axis.line=element_blank(), axis.text=element_blank(), axis.ticks=element_blank(), axis.title=element_text(size=10))+
    geom_hline(aes(yintercept=0), lty='dashed')+
    geom_vline(aes(xintercept=0), lty='dashed')+
    labs(x='MDS1 [32%]', y='MDS2 [12%]')

site.sp.plot<-site.plot+
  geom_segment(data = plant.arrow, aes(x = 0, xend = MDS1*as.numeric(rsq)*.7, y = 0, yend = MDS2*as.numeric(rsq)*.7),
               arrow = arrow(length = unit(0.2, "cm")), colour = "black")+
  geom_text(data=plant.arrow, aes(x=MDS1*as.numeric(rsq)*.7, y=MDS2*as.numeric(rsq)*.7, label=order))+xlim(c(-.6,.7))+ylim(c(-.4,.4))

ggsave(site.sp.plot, file='Fig_uncon_gen_biplot.pdf', width=84, height=90, units='mm')
```

## Constrained ordination (plant traits)  
Using all traits except total sesquiterpenes to constrain the ordinations for genetic and sex variation using dbRDA. For genetic variation use a partial dbRDA to contorl for plant sex. traits are rescaled to mean 0, sd 1.   

### Gradient analysis - 
Ordinates community data according to their similarity in species composition. Relationships between pecies and environmental gradients are investigated by correlated ordination scores with environmental variables. (vs direct which ordinates on environment and then looks at how species relate to ordination scores). Constrained uses both daasets to infer patterns in composition from patterns in environmental variables. Identifies which environmental variables are most important in structuring the community and how environmental variables are related, and how the community varies along these gradients.  


#### RDA - find linear combinations of predictor variables to represent as much variance in the response as possible. Detects gradient which all species are positively correlated, but CCA will not. 
Can do a partial RDA - can use block as covariable to factor out local effects. In RDA variance explained is easier to aprtition and interpret eignevalues. Constrained PCA (inappropriate under unimodal model). Species and environmental variables represented using arrows. 

#### dbRDA - RDA with non euclidean distances, but still linear and metric. Constrained version of metric scaling (PCoA), which uses non-Euclidean distances.   

Genetic variation:  
analysis is restricted to plants that have arthropods AND traits measured (n = 122; ~50% of plants)
```{r}
# community df with only plants that have all traits measured as well as arthropod relative density
plant.df<-cbind(arth.plant$comm.df%>%dplyr::select(ID, SEX, FAM, ROW, COL, BLOCK, WATER, RGR, flowers_all, SLA, CN, Per_H2O, MT_sum,leaf.toughness), arth.plant$comm.mat)%>%
  filter(complete.cases(MT_sum), complete.cases(Per_H2O), complete.cases(CN))

##redo matrix of arthropod relative density to correspond
plant.mat<-plant.df%>%dplyr::select(ACARINA:`NA`)%>%mutate(total=rowSums(.,na.rm=TRUE))
plant.dist<-vegdist(plant.mat, method='bray')##relative densities, bray cutus

#traits
trait.scale<-as.data.frame(scale(plant.df%>%transform(flowers_all = log(flowers_all+2), MT_sum = log(MT_sum+1))%>%dplyr::select(RGR:leaf.toughness)))%>%##scale traits
  mutate(ID=plant.df$ID)%>%left_join(plant.df%>%dplyr::select(ID:WATER), by='ID') ##add ing rouping vars

##dbRDA - conditioned on ROW (pRDA)
db.ind<-capscale(plant.dist ~ RGR+flowers_all+SLA+CN+Per_H2O+MT_sum+leaf.toughness+Condition(trait.scale$SEX), data=trait.scale, comm = plant.mat, na.action = 'na.omit')

head(summary(db.ind)) ##constrained = .11; unconstrained = .62
anova(db.ind, permutations=1000)  
anova(db.ind, by='axis', permutations=1000)
anova(db.ind, by='marginal', permutations=1000)
##traits = flowers, sla, cn
##18% explained by traits when conditioned by sex and block (residual variation not including the ocnditioned)
#CAP1 = 

## so use marginal tests? 
```  
Partial db-RDA (controlling for plant sex) on genetic variation in arthropods constrained by traits (scaled).  
#### Constrained ordination plots  
Genotype centroids:  
```{r}
#dbRDA##build the plot
#fit genotype, sex centroids, order significance
plant.env.fam<-envfit(db.ind, plant.df)
plant.env.fam
#plot centroids for each genotype
plant.fam.df<-data.frame(FAM = rownames(plant.env.fam$factors$centroids), CAP1 = plant.env.fam$factors$centroids[,1], CAP2 = plant.env.fam$factors$centroids[,2])%>%
  filter(grepl('FAM', FAM))%>%
  mutate(FAM =gsub('FAM','', FAM))%>%
  left_join(plant.df%>%dplyr::select(FAM,SEX)%>%unique()%>%transform(FAM=as.character(FAM)))

##sex centroids
plant.sex.df<-data.frame(SEX = rownames(plant.env.fam$factors$centroids), CAP1 = plant.env.fam$factors$centroids[,1], CAP2 = plant.env.fam$factors$centroids[,2])%>%
  filter(grepl('SEX', SEX))%>%
  mutate(SEX =gsub('SEX','', SEX))

##individual plants
plant.id.df<-data.frame(SEX= plant.df$SEX, ID=rownames(db.ind$CA$u),
                        FAM=plant.df$FAM, 
                        CAP1 = scores(db.ind, scaling=2, correlation=TRUE)$sites[,1], 
                        CAP2 = scores(db.ind, scaling=2, correlation=TRUE)$sites[,2])

# arrows for orders, filtered by significance and rsq
plant.arrow.sp<-data.frame(order= rownames(plant.env.fam$vectors$arrows), CAP1 = plant.env.fam$vectors$arrows[,1], 
                           CAP2 = plant.env.fam$vectors$arrows[,2], pval = plant.env.fam$vectors$pvals, rsq = plant.env.fam$vectors$r)%>%
  filter(pval <= 0.05 ,rsq > 0.10)%>%filter(order != 'RGR', order != 'flowers_all')

##ARANEAE, COLEOPTERA,DIPTERA,HEMIPTERA,HYMENOPTERA,PSOCOPTERA,THYSANOPTERA

## arrows based on species scores (not envfit) - not useing
plant.arrow.spscore<-data.frame(order=rownames(scores(db.ind, scaling=2, correlation=TRUE)$species), 
                       CAP1 = scores(db.ind, scaling=2, correlation=TRUE)$species[,1], 
                       CAP2 = scores(db.ind, scaling=2, correlation=TRUE)$species[,2])%>%filter(order %in% carrow.df$order)

##arrows for trait contraints
plant.arrow.trait<-as.data.frame(summary(db.ind)$biplot)%>%mutate(trait=rownames(.))%>%filter(trait %in% c('flowers_all','SLA','CN','Per_H2O'))
```
```{r, fig.height=8, fig.width=8}
##build troplot for constrianed ordinaton on geno
plant.triplot<-ggplot(plant.fam.df, aes(x=CAP1, y=CAP2))+
  geom_jitter(aes(fill=SEX), shape=21,size=1)+
  theme_nell()+
  scale_fill_manual(values=c('black','white'))+
  theme(axis.line=element_blank(), axis.ticks=element_blank(), axis.text=element_blank(), axis.title = element_text(size=10))+
  geom_hline(aes(yintercept=0), lty='dashed')+
  geom_vline(aes(xintercept=0), lty='dashed')+
  geom_segment(data = plant.arrow.sp, aes(x = 0, xend = CAP1*rsq*1.4, y = 0, yend = CAP2*rsq*1.4),##arrows for species scores
                                         arrow = arrow(length = unit(0.2, "cm")), colour = "black", size=1)+
  geom_text(data=plant.arrow.sp, aes(x=CAP1*rsq*1.4, y=CAP2*rsq*1.4, label=order))+
  geom_segment(data =plant.arrow.trait, aes(x = 0, xend = CAP1*1.5, y = 0, yend = CAP2*1.5), ###arrows for traits
                                         arrow = arrow(length = unit(0.2, "cm")), colour = "green", size=1)+
  geom_text(data=plant.arrow.trait, aes(x=CAP1*1.5, y=CAP2*1.5, label=trait))+
  labs(x='CAP1 [65%]', y='CAP2 [18%]')+xlim(c(-1.35,1.35))+ylim(c(-1,1))
plant.triplot #constrained ordination - genetic variation controlling for sex effects

```
### Sexual dimorphism in arhtropod composition  

```{r}
##read data
arth.geno<-list(comm.df = read.csv('/Users/colleennell/Dropbox/Projects/BASA/GENO_LSM_ARTH_ORD_COMM.csv')%>%
                  left_join(lsm.cast%>%dplyr::select(FAM,SEX,RGR,flowers_all, SLA,CN,MT_sum,ST_sum,Per_H2O, leaf.toughness)), ##add trait data
                comm.mat = read.csv('/Users/colleennell/Dropbox/Projects/BASA/GENO_LSM_ARTH_ORD_MAT.csv'))
arth.geno$comm.dist <-vegdist(arth.geno$comm.mat, method='bray')
#permanova
arth.ad.rg<-adonis(arth.geno$comm.dist~SEX, data=arth.geno$comm.df, permutations=10000)
arth.ad.rg
```  

```{r}
##same constrained analysis but using genotype means
# community df with only plants that have all traits measured as well as arthropod relative density
geno.df<-arth.geno$comm.df
geno.mat<-arth.geno$comm.mat
geno.dist<-arth.geno$comm.dist##relative densities, bray cutus

#scale traits
geno.trait.scale<-as.data.frame(scale(geno.df%>%transform(flowers_all = log(flowers_all+2), MT_sum = log(MT_sum+1))%>%dplyr::select(RGR:leaf.toughness)))%>%##scale traits
  mutate(FAM=geno.df$FAM, SEX= geno.df$SEX) ##add ing rouping vars

##dbRDA - conditioned on ROW (pRDA)
db.geno<-capscale(geno.dist ~ RGR+flowers_all+SLA+CN+Per_H2O+MT_sum+leaf.toughness, data=geno.trait.scale, comm = geno.mat, na.action = 'na.omit')

head(summary(db.geno)) ##constrained = .11; unconstrained = .62
anova(db.geno, permutations=1000)  
anova(db.geno, by='axis', permutations=1000)
anova(db.geno, by='margin', permutations=1000)
##traits = rgr, flowers, sla, pwc, 


```
```{r}
#fit genotype, sex centroids, order significance
geno.env.fam<-envfit(db.geno, geno.df)
geno.env.fam

##sex centroids
geno.sex.df<-data.frame(SEX = rownames(geno.env.fam$factors$centroids), CAP1 = geno.env.fam$factors$centroids[,1], 
                        CAP2 = geno.env.fam$factors$centroids[,2])%>%
  filter(grepl('SEX', SEX))%>%
  mutate(SEX =gsub('SEX','', SEX))

##individual genotype
geno.id.df<-data.frame(SEX= geno.df$SEX, FAM=rownames(db.geno$CA$u),
                        FAM=geno.df$FAM, 
                        CAP1 = scores(db.geno, scaling=2, correlation=TRUE)$sites[,1], 
                        CAP2 = scores(db.geno, scaling=2, correlation=TRUE)$sites[,2])

# arrows for orders, filtered by significance and rsq
geno.arrow.sp<-data.frame(order= rownames(geno.env.fam$vectors$arrows), CAP1 = geno.env.fam$vectors$arrows[,1], 
                           CAP2 = geno.env.fam$vectors$arrows[,2], pval = geno.env.fam$vectors$pvals, rsq = geno.env.fam$vectors$r)%>%
  filter(pval <= 0.05 ,rsq > 0.10)%>%filter(order != 'RGR', order != 'flowers_all', order != 'FAM')

##ARANEAE, COLEOPTERA,DIPTERA,HEMIPTERA,HYMENOPTERA,PSOCOPTERA,THYSANOPTERA

## arrows based on species scores (not envfit) - not useing
geno.arrow.spscore<-data.frame(order=rownames(scores(db.geno, scaling=2, correlation=TRUE)$species), 
                       CAP1 = scores(db.geno, scaling=2, correlation=TRUE)$species[,1], 
                       CAP2 = scores(db.geno, scaling=2, correlation=TRUE)$species[,2])%>%filter(order %in% geno.arrow.sp$order)

##arrows for trait contraints
geno.arrow.trait<-as.data.frame(summary(db.geno)$biplot)%>%mutate(trait=rownames(.))%>%
  filter(trait %in% c('flowers_all','SLA','RGR','Per_H2O'))


##build troplot for constrianed ordinaton on geno
geno.triplot<-ggplot(geno.id.df, aes(x=CAP1, y=CAP2))+
  geom_jitter(aes(fill=SEX), shape=21,size=1)+
  theme_nell()+
  labs(x='CAP1 [62%]', y='CAP2 [19%]')+
  scale_fill_manual(values=c('black','white'))+
  theme(axis.line=element_blank(), axis.ticks=element_blank(), axis.text=element_blank(), axis.title = element_text(size=10))+
  geom_hline(aes(yintercept=0), lty='dashed')+
  geom_vline(aes(xintercept=0), lty='dashed')+
  geom_segment(data = geno.arrow.sp, aes(x = 0, xend = CAP1*rsq*3, y = 0, yend = CAP2*rsq*3),##arrows for species scores
                                         arrow = arrow(length = unit(0.2, "cm")), colour = "black", size=1)+
  geom_text(data=geno.arrow.sp, aes(x=CAP1*rsq*3, y=CAP2*rsq*3, label=order))+
  geom_segment(data =geno.arrow.trait, aes(x = 0, xend = CAP1*1, y = 0, yend = CAP2*1), ###arrows for traits
                                         arrow = arrow(length = unit(0.2, "cm")), colour = "green", size=1)+
  geom_text(data=geno.arrow.trait, aes(x=CAP1*1.3, y=CAP2*1.3, label=trait))+
  xlim(c(-1.5,1.6))+ylim(c(-1.4,1.1))+
  geom_text(data=geno.sex.df, aes(x=CAP1, y=CAP2, label=SEX), size=5)+ ##sex centroids
  stat_ellipse(data=geno.id.df,aes(group=SEX, lty=SEX),type='t', level=.95) ##elolipses
  
geno.triplot #constrained ordination - genetic variation controlling for sex effects
```
```{r, fig.width=6, fig.height=10}

triplot.fig<-plot_grid(plant.triplot, geno.triplot,labels=c('(a)','(b)'), nrow=2, ncol=1)
triplot.fig

ggsave(triplot.fig, file='Fig_uncon_gen_triplot.pdf', width=100, height=210,units='mm')
```

```{r}
pco.geno<-capscale(arth.geno$comm.dist ~ 1 , data=arth.geno$comm.df, comm = arth.geno$comm.mat)
#str(arth.geno$comm.df)
eig<-eigenvals(pco.geno)##extract eignevalues
eig[1:3] / sum(eig) #proportion of variance explained by each axis
cumsum(eig[1:3] / sum(eig))

head(summary(pco.geno))
```  
  
#### Unconstrained ordination of plant genotypes with centroids for sex  
```{r}
plot.g<-plot_cap_df(pco.geno, species=TRUE, data=arth.geno$comm.df, scaly=2)
plot.gen<-unc_plot(df=plot.g, fam.df=plot.g$ind.df, ind.df = plot.g$ind.df,species=plot.g$sp.df, env=plot.g$sp.env)

###########
#plot centroids for each genotype, unconstrained ordination
geno.ind<-data.frame(ID = rownames(pco.geno$CA$u), SEX= arth.geno$comm.df$SEX, 
                        FAM=arth.geno$comm.df$FAM,
                        MDS1 = scores(pco.geno, scaling=2, correlation=TRUE)$sites[,1], 
                        MDS2 = scores(pco.geno, scaling=2, correlation=TRUE)$sites[,2])

##this si scaling 2 (species axes correlations)
geno.sp<-data.frame(label=rownames(scores(pco.geno, scaling=2, correlation=TRUE)$species), 
                       MDS1 = scores(pco.geno, scaling=2, correlation=TRUE)$species[,1], 
                       MDS2 = scores(pco.geno, scaling=2, correlation=TRUE)$species[,2])

##use envfit for fitted order arrows
#inertia is variance in species abundance
bt.geno<-cbind(arth.geno$comm.df%>%dplyr::select(SEX,FAM), arth.geno$comm.mat)
pco.env.geno<-envfit(pco.geno, bt.geno)
pco.env.geno
sex.geno<-data.frame(FAM = rownames(pco.env.geno$factors$centroids), MDS1 = pco.env.geno$factors$centroids[,1], MDS2 = pco.env.geno$factors$centroids[,2]))%>%filter(FAM == 'SEXM' | FAM =='SEXF')

geno.arrow<-data.frame(order= rownames(pco.env.geno$vectors$arrows), MDS1 = pco.env.geno$vectors$arrows[,1], MDS2 = pco.env.geno$vectors$arrows[,2], pval = pco.env.geno$vectors$pvals, rsq = pco.env.geno$vectors$r)%>%
  filter(pval <= 0.05, rsq > 0.1)

##axis labels with var explained
xlabel<-paste0('MDS1 (', round(100*cumsum(eig / sum(eig))[[1]],2), '% of total variation)')
ylabel<-paste0('MDS2 (', round(100*(eig[[2]] / sum(eig)),2), '% of total variation)')

###build the plot
site.plot.geno<-ggplot(geno.ind, aes(x=MDS1, y=MDS2))+
    geom_jitter(aes(fill=SEX), shape=21,size=1.5)+
    theme_nell()+
    scale_fill_manual(values=c('black','white'))+
    theme(axis.line=element_blank())+
    geom_hline(aes(yintercept=0), lty='dashed')+
    geom_vline(aes(xintercept=0), lty='dashed')+
    labs(x='MDS1 [46%]', y='MDS2 [15%]')

site.sp.plot.geno<-site.plot.geno+
  geom_segment(data = plant.arrow.geno, aes(x = 0, xend = MDS1*as.numeric(rsq)*.5, y = 0, yend = MDS2*as.numeric(rsq)*.5),
               arrow = arrow(length = unit(0.2, "cm")), colour = "black", size=1)+
  geom_text(data=plant.arrow.geno, aes(x=MDS1*as.numeric(rsq)*.5, y=MDS2*as.numeric(rsq)*.5, label=order))+
  theme(axis.line=element_blank(), axis.text=element_blank(), axis.ticks=element_blank(), axis.title=element_text(size=10))

##add centroids for each sex and 95% CI
sex.plot.geno<-site.sp.plot.geno+
  geom_text(data=sex.center.geno, aes(x=MDS1, y=MDS2, label=FAM), size=5)+
  stat_ellipse(aes(group=SEX, lty=SEX),type='t', level=.95)+
  scale_fill_manual(values=c('black','white'))+theme(legend.position='none')+xlim(c(-.9,1))+ylim(c(-.6,.4))


###build the plot
##a plot
site.plot<-ggplot(fam.center, aes(x=MDS1, y=MDS2))+
    geom_jitter(aes(fill=SEX), shape=21,size=1.5)+
    theme_nell()+
    scale_fill_manual(values=c('black','white'))+
    theme(axis.line=element_blank(), axis.text=element_blank(), axis.ticks=element_blank(), axis.title=element_text(size=10))+
    geom_hline(aes(yintercept=0), lty='dashed')+
    geom_vline(aes(xintercept=0), lty='dashed')+
    labs(x='MDS1 [32%]', y='MDS2 [12%]')


site.sp.plot<-site.plot+
  geom_segment(data = plant.arrow, aes(x = 0, xend = MDS1*as.numeric(rsq)*.7, y = 0, yend = MDS2*as.numeric(rsq)*.7),
               arrow = arrow(length = unit(0.2, "cm")), colour = "black")+
  geom_text(data=plant.arrow, aes(x=MDS1*as.numeric(rsq)*.7, y=MDS2*as.numeric(rsq)*.7, label=order))+xlim(c(-.6,.7))+ylim(c(-.4,.4))

#ggsave(site.sp.plot, file='Fig_uncon_gen_biplot.pdf', width=84, height=90, units='mm')

uncon.ords<-plot_grid(site.sp.plot,sex.plot.geno, nrow=2, ncol=1, labels=c('(a)','(b)'))
uncon.ords
ggsave(uncon.ords, file='uncon.ords.biplot.pdf', width=100, height=210, units='mm')
```
